The Analogue Method for Precipitation Prediction: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2016-12-01 Moving temporal window - HESS/template.tex
Elsevier edition.

@@

%TODO: english editing
%TODO: perspective Obled: MTW without precip -> indicator

@@

\title{The Analog Method for Precipitation Prediction: Finding Better Analog Situations at a Sub-Daily Time Step}

@@

\runningtitle{Finding Better Analog Situations at a Sub-Daily Time Step}

@@

\begin{abstract}
The Analog Method (AM) aims at predicting local weather variables (predictands), such as precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally first assessed on gradients of geopotential fields, in order to sample days with a similar atmospheric circulation. Other predictors, such as moisture variables, can also be added in a successive level of analogy.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The main reason is the use of daily precipitation time series, due to the length of their available archives, and the unavailability of equivalent archives at a finer time step. However, it is unlikely for the best analogy to happen at the very same hour, as it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog rank due to an accumulation of more similar situations in the selection. A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to stronger diurnal cycles in summer, which favor predictors at the same hour for the target and analog days.

The impact of the MTW on the prediction skill has been assessed by means of a sub-daily precipitation series transformed into moving 24~h-totals at a 6-hourly time step. The prediction skill was found to be improved by the MTW, and even more after recalibrating the AM parameters. Moreover, the improvement was greater for days with heavy precipitation, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing archives of measurements, or by the establishment of regional precipitation reanalyzes. Then, the use of a MTW in AMs can be considered for any application, and in any context, let it be operational forecasting or climate-related studies.

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of statistical downscaling methods \citep{Maraun2010}, and consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the method exist \citep[a non-exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method is analyzed in \citet{Chardon2014}. A great advantage of AMs is that they create realistic precipitation patterns, as they are based on observed situations, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} have analyzed the influence of the archive length on the quality of the analogy. They highlight a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (or degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrate that it is not possible to find good analogs on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to search for analogs, and the archive length is maximized.

Thus, due to the availability of long precipitation archives at a daily time step that have no equivalent at a finer resolution, AMs are usually implemented on a daily basis (see references above). Therefore, the analog situations are assessed by comparing predictors at fixed hours of the day. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day. Previous tests showed the benefit, in terms of analogy criterion values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the prediction skills \citep{Finet2008}.

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant correlation between successive situations within the same day. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does extract new information from the archive. Thus, if the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to use a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new parameters \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

Section \ref{sec:data_methods} presents the context of the study as well as data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefit of introducing a MTW was assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion} and the conclusions are found in Sect. \ref{sec:conclusions}.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The data were available at an hourly time step on 25~years (1982 to 2007), and were also obtained at a daily time step (from 6~h~UTC to 6~h~UTC the next day) on 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will thereafter be presented arbitrarily for the Ulrichen station, but were equivalent for all stations.

Synoptic-scale variables, used as predictors, were extracted from the NCEP/NCAR reanalysis I dataset \citep{Kalnay1996} with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but it is not expected to affect the conclusions of the present study (see discussion in Sect. \ref{sec:old_reanalysis}).

\subsection{The considered analog method}
\label{sec:analog_method}

The first considered AM is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching for analog situations to a target day starts by a seasonal stratification through a preselection step of the potential candidates for analogy. The restriction is here a search for analog days within a 4 months window centered on the target date, for every year of the archive. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score (S1), which is a comparison of gradients, on geopotential heights over a certain spatial window and at certain hours. The smaller the S1 values, the more similar the pressure fields.

The predictor variables, extracted from reanalysis datasets, are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are thus compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected is found by \citet{Bontron2004} to have a significant influence.

Then, $N_{1}$ dates with the lowest values of S1 are considered as analogs to the target day, $N_{1}$ being a parameter to calibrate (see Sect. \ref{sec:calibration}). Finally, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day. This method will be named 2Z further on.

The second reference method (2Z-2MI, Table \ref{table:method_2Z-2MI}) adds a subsequent level of analogy on moisture variables, compared by means of the root-mean-square error (RMSE). The additional predictor is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled from the $N_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations. When a second level of analogy is added, a higher number of $N_{1}$ analogs is kept on the first level.

More complex AMs exist, with additional predictors \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. The MTW can also be applied on these, as they generally rely on a similarity of the atmospheric circulation in the first level of analogy, but it is easier to interpret the impact of the MTW on more basic methods.

@@

In order to assess the performance of AMs, the continuous ranked probability score \citep[CRPS,][]{Brown1974, Matheson1976, Hersbach2000} is often employed \citep[see, e.g.,][]{Bontron2004, Bontron2005, BenDaoud2008, Horton2012, Marty2012, Radanovics2013, Chardon2014, Junk2015, BenDaoud2016, Caillouet2016}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. It is defined as follows:

\begin{equation}
\label{eq:CRPS}
S_{\text{CRP}} = \int_{-\infty}^{+\infty} \left[ F(y)-\text{H}(y-y^{0})\right]^{2} \text{d}y ,  
\end{equation}
where $\text{H}(y-y^{0})$ is the Heaviside function that is null when $y-y^{0}<0$, and has the value 1 otherwise.

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution (i.e. the distribution of all precipitation values from the corresponding archive) as the reference. The continuous ranked probability skill score (CRPSS) is thus defined as follows:

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ would be the one for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS: $S_{\text{SCRP}}~=~1$ for a perfect prediction, and $S_{\text{SCRP}}~<~0$ for a prediction with a lower skill than the reference.

@@

The moving time window (MTW) technique aims at finding better analog situations at different hours of the day rather than comparing the predictors at fixed hours. Thus, instead of looking for analogs at a 24-hourly time step, they are sought at the time step matching the predictor temporal resolution, which is here 6-hourly (Fig. \ref{fig:principle}).

The target situations and their corresponding observed precipitation values do not change, because the prediction is still established on a daily basis, for a fixed period of the target day (6~h~UTC to 6~h~UTC the next day), as before. The difference is that the candidates are 4~times more numerous (even though they are not fully independent). No constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

In order to assess the benefit of searching analog situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high temporal resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step, by means of a moving 24~h-total.  As sub-daily precipitation time series are available on a shorter period than daily ones, the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007) needs to be assessed.

@@

It would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008), rather than being limited to the reduced period corresponding to the sub-daily precipitation data (cf. previous section). Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily precipitation time series were assessed.

The first technique was a proportional distribution, where the observed daily precipitations were considered constant during the measuring period (6~h~UTC to 6~h~UTC the next day). It then consisted in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}).

The second approach aimed at getting closer to the chronology of the actual precipitation by relying on some informative variables during the reconstruction procedure. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) were used in this perspective, despite their rough resolution. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Then, once the best proxy has been selected, its temporal profile was used in order to disaggregate the daily precipitation time series.

\subsection{Calibration of the analog method}
\label{sec:calibration}

The semi-automatic sequential procedure elaborated by \citet{Bontron2004} was used to calibrate the AM \cite[see also ][]{Radanovics2013, BenDaoud2016}. The analogy levels (e.g. the atmospheric circulation or moisture index) are calibrated sequentially. The parameters calibrated by this approach, for every level of analogy, are the spatial windows on which the predictors are compared and the number of analogs. The procedure consists of the following steps \citep{Bontron2004}:

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable (e.g. geopotential height, temperature, relative humidity, etc.),
		\item pressure level (e.g. 500 hPa, ...),
		\item temporal window (hour of the day \textendash\ e.g. 6, or 12~h~UTC),
		\item initial analog numbers (e.g. $N_{1}=50$).
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criterion) over a large domain. Every point (or cell) of the full domain is assessed on all predictors of the level of analogy, jointly (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analog situations is then reconsidered and optimized for the current level of analogy.
	\end{enumerate}
	\item A new level of analogy can then be added, based on other variables (such as the moisture index) at chosen pressure levels and hours of the day. The number of analogs for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (2) for the new level. The parameters calibrated on the previous analog levels are fixed and do not change. 
	\item Finally, the numbers of analogs for the different levels of analogy are re-assessed. This is done iteratively by varying the number of analogs of each level in a systematic way.
\end{enumerate}

The calibration is thus done in successive steps, on a limited number of parameters. Previously calibrated parameters are generally not reassessed (except for the number of analogs). More advanced techniques, such as using genetic algorithms \citep{Horton2016}, exist, but are out of the scope of the present study.

@@

The performance loss resulting from the reduction of a 47~years archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series, see Sect. \ref{sec:mtw}) was assessed with the original method, without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analog numbers in Tables \ref{table:analog_nb}) and were used thereafter.

The impact of the archive reduction can be seen in Fig. \ref{fig:plots_CRPSS_2Z} for the 2Z method and in Fig. \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z-2MI method. As expected, a loss of performance could be observed for each station, except for that of Aigle, which seemed relatively indifferent to this change. This loss was globally significant, with up to -1.89 points for Visp and the 2Z method.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons were substantial, with greater improvements for winter than summer.

@@

\subsection{Improvement of the selection of analog situations}

>>>>

%TODO: check references: complete, doi, etc

@@

@@

	\title{The Analogue Method for Precipitation Prediction: Finding Better Analogue Situations at a Sub-Daily Time Step}

@@

	\runningtitle{Finding Better Analogue Situations at a Sub-Daily Time Step}

@@

	\begin{abstract}
		Analogue methods (AM) predict local weather variables (predictands), such as precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally assessed on gradients of geopotential fields first, in order to sample days with a similar atmospheric circulation. Other predictors, such as moisture variables, can also be added in a successive level of analogy.

		The search for candidate situations for a given target day is usually undertaken by comparing the state of the atmosphere at fixed hours of the day for both the target day and the candidate analogues. The main reason is the use of daily precipitation time series due to the length of their available archives, and the unavailability of equivalent archives at a finer time step. However, it is unlikely for the best analogy to occur at the very same hour, as it may occur at a different time of day. In order to assess the potential for finding better analogues at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

		The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the entire distribution of analogue dates. The improvement was found to grow with the analogue rank due to an accumulation of better analogues in the selection. A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to stronger diurnal cycles in summer that favour predictors at the same hour for the target and analogue days.

		The impact of the MTW on prediction skill has been assessed by means of a sub-daily precipitation series transformed into moving 24-h totals at 6-h time steps. The prediction skill was found to have improved by the MTW, and even to a greater extent after recalibrating the AM parameters. Moreover, the improvement was greater for days with heavy precipitation, which are generally related to more dynamic atmospheric situations where timing is more specific and which are fewer in the archive.

		However, the MTW requires a sub-daily precipitation time series, which is generally available for a shorter period than daily precipitation archives. Therefore, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, emphasizing the need to use time series with an appropriate chronology. These should be available in the near future, either by means of growing archives of measurements or by the establishment of regional precipitation reanalysis data. The use of an MTW in the AM could then be considered for any application and in any context, may it be for operational forecasting or climate-related studies.

@@

	Analogue methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are statistical downscaling methods \citep{Maraun2010} and consist of finding past situations that are similar in terms of the atmospheric circulation or other synoptic predictors to the target day of interest. The local weather variables of interest (predictand) that were observed at the analogue dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the method exist \citep[a non-exhaustive listing can be found in][]{BenDaoud2016}. The AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as in precipitation downscaling from a climate perspective \citep[e.g.][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

	The spatial transferability of the method is analysed in \citet{Chardon2014}. A great advantage of AMs is that they create realistic precipitation patterns, since they are based on observed situations with consistent spatial distribution, as long as the analogue dates chosen for a region are the same. Their temporal transferability has also been the topic of recent studies for past or future climates \citep{Dayon2015, Caillouet2016}.

	\citet{Ruosteenoja1988} and \citet{Vandendool1994} have analysed the influence of the archive length on the quality of the analogy. They highlighted a three-way relationship between the quality of the analogy, the archive length, and the size of the spatial domain (or degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrated that it is not possible to find good analogues over the whole northern hemisphere with a 100 year archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered when searching for analogues, and the archive length is maximized.

	Therefore, due to the availability of long precipitation archives at a daily time step that have no equivalent at a finer resolution, AMs are usually implemented on a daily basis. Therefore, the analogue situations are assessed by comparing predictors at fixed hours of the day. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day and that better candidates can be found by shifting to a different hour. With this assumption, a moving time window (MTW) was introduced to allow the search for candidates at different hours of the day. Previous tests showed the benefit, in terms of analogy criterion values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the prediction skill \citep{Finet2008}.

	The MTW finds analogue situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive due to significant correlation between successive situations within the same day. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does extract new information from the archive. Therefore, if the reduction of the archive length needed by the MTW, due to the reduced availability of a sub-daily precipitation time series, is expected to decrease the AM performance, the archive inflation brought by the MTW is expected to contribute to an increase in performance.

	Other possibilities exist for increasing the prediction skill of the AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to use a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new parameters \citep{Horton2016}. However, the MTW technique is not in competition with other approaches and can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

	Section \ref{sec:data_methods} presents the context of the study as well as the data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefits of introducing a MTW were assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}) and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion}, and the conclusions are found in Sect. \ref{sec:conclusions}.

@@

	The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from six automatic weather stations, viz., Ulrichen, Zermatt, Visp, Montana, Sion, and Aigle (Fig. \ref{fig:map}) that are subject to various meteorological influences \citep{Horton2012}. The data were available at an hourly time step for 25~years (1982 to 2007) and were also obtained at a daily time step (from 6~h~UTC to 6~h~UTC the next day) for 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will hereafter be presented arbitrarily for the Ulrichen station but are equivalent for all stations.

	Synoptic-scale variables, used as predictors, were extracted from the NCEP/NCAR reanalysis I dataset \citep{Kalnay1996} with a 6-hour temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but it is not expected to affect the conclusions of the present study (see discussion in Sect. \ref{sec:old_reanalysis}).

	\subsection{The considered analogue method}
	\label{sec:analog_method}

	The first considered AM is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching for analogue situations to a target day starts by a seasonal stratification through a preselection step of the potential candidates for analogy. The restriction is a search for analogue days within a 4-month window centred on the target date for every year of the archive. The similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet{Teweles1954} score (S1) that is a comparison of gradients on geopotential heights over a certain spatial window and at certain hours. With smaller S1 values, there is greater similarity in the pressure fields.

	The predictor variables extracted from reanalysis datasets are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected is found by \citet{Bontron2004} to have a significant influence.

	Then, $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day, $N_{1}$ being a parameter to calibrate (see Sect. \ref{sec:calibration}). Finally, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic prediction for the target day. This method will be named 2Z.

	The second reference method (2Z-2MI, Table \ref{table:method_2Z-2MI}) adds a subsequent level of analogy with moisture variables, compared by means of the root-mean-square error (RMSE). The additional predictor is a moisture index composed of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled from the $N_{1}$ analogues on the atmospheric circulation, resulting in a smaller number of analogue situations. When a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

	More complex AMs exist with additional predictors \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. The MTW can also be applied to these, as they generally rely on a similarity of the atmospheric circulation in the first level of analogy. However, it is easier to interpret the impact of the MTW using more basic methods.

@@

	In order to assess the performance of AMs, the continuous ranked probability score \citep[CRPS,][]{Brown1974, Matheson1976, Hersbach2000} is often employed \citep[see, e.g.,][]{Bontron2004, Bontron2005, BenDaoud2008, Horton2012, Marty2012, Radanovics2013, Chardon2014, Junk2015, BenDaoud2016, Caillouet2016}. It allows the evaluation of the predicted cumulative distribution functions $F(y)$ of the precipitation values $y$ from analogue situations compared to the observed value $y^{0}$. A better prediction has a smaller score. It is defined as follows:

	\begin{equation}
	\label{eq:CRPS}
	S_{\text{CRP}} = \int_{-\infty}^{+\infty} \left[ F(y)-\text{H}(y-y^{0})\right]^{2} \text{d}y ,  
	\end{equation}
	where $\text{H}(y-y^{0})$ is the Heaviside function that is null when $y-y^{0}<0$ and has the value of 1 otherwise.

	In order to compare the value of the score in regard to a reference, one often considers its skill score expression and uses the climatological distribution (i.e., the distribution of all precipitation values from the corresponding archive) as the reference. The continuous ranked probability skill score (CRPSS) is defined as follows:

	\begin{equation}
	\label{eq:CRPSS}
	S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
	\end{equation}
	where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ is for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS: $S_{\text{SCRP}}~=~1$ for a perfect prediction and $S_{\text{SCRP}}~<~0$ for a prediction with a lower skill than the reference.

@@

	The moving time window (MTW) technique aims at finding better analogue situations at different hours of the day rather than comparing the predictors at fixed hours. Therefore, instead of looking for analogues at a 24-hourly time step, they are sought at the time step matching the predictor temporal resolution which is a 6-hour time step in this study (Fig. \ref{fig:principle}).

	The target situations and their corresponding observed precipitation values do not change because the prediction is still established on a daily basis for a fixed period of the target day (6~h~UTC to 6~h~UTC the next day) as before. The difference is that the candidates are four times more numerous (even though they are not fully independent). No constraint was added in order to restrict the selection of multiple analogues within the same candidate dates.

	In order to assess the benefit of searching for analogue situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high temporal resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed at a 6-hour time step by means of a moving 24~h-total. Since sub-daily precipitation time series are available on a shorter period than daily ones, the loss of performance resulting from a reduction of a 47~year archive (1961 to 2008) to 25~years (1982 to 2007) needs to be assessed.

@@

	It would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008), rather than being limited to the reduced period corresponding to the sub-daily precipitation data (cf. previous section). Therefore, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily precipitation time series were assessed.

	The first technique was a proportional distribution, where the observed daily precipitations were considered constant during the measuring period (6~h~UTC to 6~h~UTC the next day). Proportional parts of the original daily time series were allocated into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}).

	The second approach aimed at getting closer to the chronology of the actual precipitation by relying on some informative variables during the reconstruction procedure. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) were used for this purpose, despite their rough resolution. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and a time lapse between both series was allowed to take into account the significant distance separating the weather stations and the reanalysis grid point. The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series over the period 1982-2007. Once the best proxy had been selected, its temporal profile was used in order to disaggregate the daily precipitation time series.

	\subsection{Calibration of the analogue method}
	\label{sec:calibration}

	The semi-automatic sequential procedure elaborated by \citet{Bontron2004} was used to calibrate the AM \cite[see also ][]{Radanovics2013, BenDaoud2016}. The analogy levels (e.g. the atmospheric circulation or moisture index) are calibrated sequentially. The parameters calibrated by this approach, for every level of analogy, are the spatial windows on which the predictors are compared and the number of analogues. The procedure consists of the following steps \citep{Bontron2004}:

	\begin{enumerate}
		\item Manual selection of the following parameters:
		\begin{enumerate}
			\item meteorological variable (e.g. geopotential height, temperature, relative humidity, etc.),
			\item pressure level (e.g. 500 hPa, ...),
			\item temporal window (hour of the day \textendash\ e.g. 6 or 12~h~UTC),
			\item initial analogue numbers (e.g. $N_{1}=50$).
		\end{enumerate}

		\item For every level of analogy:
		\begin{enumerate}
			\item Identification of the most skilled unitary cell (one point for moisture variables and four for the geopotential height when using the S1 criterion) over a large domain. Every point (or cell) of the full domain is assessed on all predictors of the level of analogy, jointly (consisting generally of the same variable but on different pressure levels and at different hours).
			\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The spatial window grows by repeating the previous steps until no improvement is reached.
			\item The number of analogue situations is then reconsidered and optimized for the current level of analogy.
		\end{enumerate}
		\item A new level of analogy can then be added, based on other variables (such as the moisture index) at chosen pressure levels and hours of the day. The number of analogues for the next level of analogy is initiated at a chosen value. The procedure starts again from step 2 (calibration of the spatial window and the number of analogues) for the new level. The parameters calibrated on the previous analogue levels are fixed and do not change. 
		\item Finally, the numbers of analogues for the different levels of analogy are reassessed. This is done iteratively by varying the number of analogues of each level in a systematic way.
	\end{enumerate}

	The calibration is done in successive steps with a limited number of parameters. Previously calibrated parameters are generally not reassessed (except for the number of analogues). More advanced techniques, such as using genetic algorithms \citep{Horton2016}, exist but are out of the scope of the present study.

@@

	The performance loss resulting from the reduction of a 47~year archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series in Sect. \ref{sec:mtw}) was assessed with the original method without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analogue numbers in Tables \ref{table:analog_nb}) and were used thereafter.

	The impact of the archive reduction can be seen in Fig. \ref{fig:plots_CRPSS_2Z} for the 2Z method and in Fig. \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z-2MI method. As expected, a loss of performance was observed for each station, except for that of Aigle, which seemed relatively indifferent to this change. This loss was globally significant, with up to -1.89 points for Visp and the 2Z method.

@@

	Atmospheric dynamics vary greatly from one season to another, reflecting on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons were substantial, with greater improvements for winter than summer.

@@

	\subsection{Improvement of the selection of analogue situations}
